---
title: "Disagreement Regarding 2016 Presidential Election in the United States^[Code and data are available at: https://github.com/SANGWOONG-LEE/Thailand_Healthcare]"
subtitle: "Possible factors effecting 2016 election"
author: 
  - SangWoong Lee
  - Young Suk
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "This paper has been examined possible variables that influenced U.S citizens' choice for the 2016 U.S presidential election. We have found that there exists significantly different opinions regarding gender, wealth, race and age. This is important in a sesne that it shows thought trends as well as crucial decision indicators for voting in U.S.,With further survey at the end of this paper, we are going to explore whether this is happening only in U.S or not."
output:
  bookdown::pdf_document2:
    fig_caption: yes
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(dplyr)
library(knitr)
```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point). 



# Data
```{r, include=FALSE, echo=FALSE}
# Read the extracted and cleaned data, data extraction done in 01-gather-data.R and modification done in 02-clean-and-prepare-data.R
data <- read_csv("outputs/data/cleaned_data.csv")

urban_rural <- pivot_longer(data[1:2, 1:5], -c(Variable, Received_Immunization), names_to="immune_status", values_to="value")
region <- pivot_longer(data[3:7, 1:5], -c(Variable, Received_Immunization), names_to="immune_status", values_to="value")
education <- pivot_longer(data[8:11, 1:5], -c(Variable, Received_Immunization), names_to="immune_status", values_to="value")
religion <- pivot_longer(data[12:13, 1:5], -c(Variable, Received_Immunization), names_to="immune_status", values_to="value")
total <- pivot_longer(data[14, 1:5], -c(Variable, Received_Immunization), names_to="immune_status", values_to="value")
population <- data.frame(Variable = data$Variable, 
                         Population = data$Number_of_Children) %>% head(-1)
population$Variable <- factor(population$Variable,
                              levels = population$Variable) %>% fct_rev()
```
```{r}

# create new dataset only for urban-rural data
# urban_rural <- data.frame(location = c(rep("urban", data[1,6]), rep("rural", data[2,6])),
#                 immune_data =  c(rep("Health Report", round(data[1,2]*data[1,6]/100)), rep("Mothers Report", round(data[1,3]*data[1,6]/100)), rep("No immune", round(data[1,5]*data[1,6]/100)), 
#                                  rep("Health Report", round(data[2,2]*data[2,6]/100)), rep("Mothers Report", round(data[2,3]*data[2,6]/100)), rep("No immune", round(data[2,5]*data[2,6]/100)))
#                 
#                 )

# create new dataset only for for region
# region <- data.frame(region = c(rep("north", data[3,6]), rep("north_east", data[4,6]), rep("central", data[5,6]), rep("south", data[6,6]), rep("bangkok", data[7,6])),
#                     immune_data =  c(rep("Health Report", round(data[3,2]*data[3,6]/100)), rep("Mothers Report", round(data[3,3]*data[3,6]/100)), rep("No immune", round(data[3,5]*data[3,6]/100)), 
#                                  rep("Health Report", round(data[4,2]*data[4,6]/100)), rep("Mothers Report", round(data[4,3]*data[4,6]/100)), rep("No immune", round(data[4,5]*data[4,6]/100)),
#                                  rep("Health Report", round(data[5,2]*data[5,6]/100)), rep("Mothers Report", round(data[5,3]*data[5,6]/100)), rep("No immune", round(data[5,5]*data[5,6]/100)), 
#                                  rep("Health Report", round(data[6,2]*data[6,6]/100)), rep("Mothers Report", round(data[6,3]*data[6,6]/100)), rep("No immune", round(data[6,5]*data[6,6]/100)),
#                                  rep("Health Report", round(data[7,2]*data[7,6]/100)), rep("Mothers Report", round(data[7,3]*data[7,6]/100)), rep("No immune", round(data[7,5]*data[7,6]/100)))
#                     )

# create new dataset only for for education. due to an error in the dataset provided by DHS, some rows will be filled with NA.

 # no_edu <-data.frame(education = c(rep("no_edu", 265)),
 #                     immune_data = c(rep("Health Report", round(data[8,2]*data[8,6]/100)), rep("Mothers Report", round(data[8,3]*data[8,6]/100)), rep("No immune", round(data[8,5]*data[8,6]/100)))
 #                     )
 # 
 # primary <-data.frame(education = c(rep("primary", data[9,6])),
 #                     immune_data = c(rep("Health Report", round(data[9,2]*data[9,6]/100)), rep("Mothers Report", floor(data[9,3]*data[9,6]/100)), rep("No immune", round(data[9,5]*data[9,6]/100)))
 #                     )
 # 
 # secondary <-data.frame(education = c(rep("secondary", data[10,6])),
 #                     immune_data = c(rep("Health Report", floor(data[10,2]*data[10,6]/100)), rep("Mothers Report", floor(data[10,3]*data[10,6]/100)), rep("No immune", round(data[10,5]*data[10,6]/100)))
 #                    )
# education <- data.frame(education = c(rep("no_education", round(data[8,2]*data[8,6]/100) + round(data[8,3]*data[8,6]/100) + round(data[8,5]*data[8,6]/100)), 
#                                       rep("primary", round(data[9,2]*data[9,6]/100) + round(data[9,3]*data[9,6]/100) + round(data[9,5]*data[9,6]/100)), 
#                                       rep("secondary", round(data[10,2]*data[10,6]/100) + round(data[10,3]*data[10,6]/100) + round(data[10,5]*data[10,6]/100)),
#                                       rep("higher", round(data[11,2]*data[11,6]/100) + round(data[11,3]*data[11,6]/100) + round(data[11,5]*data[11,6]/100))),
#                     immune_data =  c(rep("Health Report", round(data[8,2]*data[8,6]/100)), rep("Mothers Report", round(data[8,3]*data[8,6]/100)), rep("No immune", round(data[8,5]*data[8,6]/100)),
#                                  rep("Health Report", round(data[9,2]*data[9,6]/100)), rep("Mothers Report", round(data[9,3]*data[9,6]/100)), rep("No immune", round(data[9,5]*data[9,6]/100)),
#                                  rep("Health Report", round(data[10,2]*data[10,6]/100)), rep("Mothers Report", round(data[10,3]*data[10,6]/100)), rep("No immune", round(data[10,5]*data[10,6]/100)),
#                                  rep("Health Report", round(data[11,2]*data[11,6]/100)), rep("Mothers Report", round(data[11,3]*data[11,6]/100)), rep("No immune", round(data[11,5]*data[11,6]/100)))
#                     )

# create new dataset only for for education
# religion <- data.frame(religion = c(rep("Buddhist", data[12,6]), rep("Islam", data[13,6])),
#                 immune_data =  c(rep("Health Report", round(data[12,2]*data[12,6]/100)), rep("Mothers Report", round(data[12,3]*data[12,6]/100)), rep("No immune", round(data[12,5]*data[12,6]/100)), 
#                                  rep("Health Report", round(data[13,2]*data[13,6]/100)), rep("Mothers Report", round(data[13,3]*data[13,6]/100)), rep("No immune", round(data[13,5]*data[13,6]/100)))
#                       )
# 
# total <- data.frame(total = c(rep("total", data[14,6])),
#                 immune_data =  c(rep("Health Report", floor(data[14,2]*data[14,6]/100)), rep("Mothers Report", floor(data[14,3]*data[14,6]/100)), rep("No immune", floor(data[14,5]*data[14,6]/100)))
#                       )
```

```{r}
population %>%
  ggplot(aes(x = Variable, y = Population, fill = Variable)) +
  geom_bar(stat='identity', position = "dodge") +
  coord_flip()
```

```{r}
#ggplot(data[1:2,], aes(x = Variable, y = Health_Record)) + geom_bar(stat='identity') + facet_wrap(~Mothers_Report)
urban_rural %>%
  ggplot(aes(x = immune_status, y = value, fill = Variable)) +
  geom_bar(stat='identity', position = "dodge") +
  coord_flip()
```
```{r}
#ggplot(data[3:7,], aes(x = Variable, y = Health_Record, fill = Variable)) + geom_bar(stat='identity')
region %>%
  ggplot(aes(x = Variable, y = value, fill = immune_status)) +
  geom_bar(stat='identity', position = "dodge") +
  coord_flip()
```
```{r}
#ggplot(data[8:11,], aes(x = Variable, y = Health_Record, fill = Variable)) + geom_bar(stat='identity')
education %>%
  ggplot(aes(x = Variable, y = value, fill = immune_status)) +
  geom_bar(stat='identity', position = "dodge") +
  coord_flip()
```
```{r}
#ggplot(data[12:13,], aes(x = Variable, y = Health_Record, fill = Variable)) + geom_bar(stat='identity')
religion %>%
  ggplot(aes(x = immune_status, y = value, fill = Variable)) +
  geom_bar(stat='identity', position = "dodge") +
  coord_flip()
```
Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r}
total %>%
  ggplot(aes(x = immune_status, y = value, fill = immune_status)) +
  geom_bar(stat='identity', position = "dodge") +
  coord_flip()
```

Talk way more about it. 



# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful. 

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional details


\newpage


# References


